\remove{Serverless workflow orchestrators make it easier for developers to
build complex serverless applications by driving the execution of a workflow
from a central service.  Making orchestrators fault-tolerant and scalable
requires engineering effort as well as large up-front resource costs to
deploy. As a result, developers typically rely on multi-tenant services run by
cloud providers that can amortize these costs across many users.

Unfortunately, this centralization comes at the expense of application
flexibility. Orchestrators support a fixed set of application patterns, and
trade off performance, scalability, fault tolerance, and resource efficiency
using a policy chosen by the cloud provider. They support many applications
well, but cannot simultaneously satisfy the needs of \emph{all} applications.

We argue that centralized serverless orchestrators are actually unnecessary.
Instead, a decentralized orchestrator, deployed as a library included by the
serverless application, can support the same application patterns and fault
tolerance as existing orchestrators using only the basic building blocks of
serverless platforms---a scalable ``Function'' executor (FaaS) and a
consistent scalable datastore. This enables different applications to use or
customize different libraries to implement different patterns or make
application-specific trade-offs.

To demonstrate this, we describe the design and implementation of \name{}, a
decentralized serverless workflow system. \name{} compiles workflows written
in existing high-level workflow languages into a partitioned orchestration
library that executes orchestration logic in-situ with user functions. \name{}
provides the same strong correctness guarantees under faults as a centralized
orchestrator, performs similarly, scales better, and costs less than existing
centralized workflow systems.

Our implementation of \name{} supports a superset of applications the AWS Step
Functions orchestrator supports, runs on both AWS Lambda and Google Cloud
Functions, and runs as much as 2x faster and up 9x cheaper than native Step
Functions on a representative set of applications.}

\majoredits{Standalone serverless orchestrators simplify the developement of
complex serverless applications but limit application flexibility and are
expensive to use. We argue that these additional orchestration services are
unnecessary. Instead, application-level orchestration, deployed as a library,
can support the same patterns and execution guarantees as existing standalone
orchestrators, using only basic serverless components that are already used by
applications and universally supported by cloud providers---a scalable
``Function'' executor (FaaS) and a consistent datastore.

To demonstrate this, we present \name{}, an application-level serverless
workflow system. \name{} provides orchestration as a library that runs in-situ
with user-defined FaaS functions. Without a standalone service, \name{}
introduces an intermediate language to partition high-level application logic
and leverages strongly consistent data stores to coordinate and ensure
correctness in a decentralized manner.

\name{} provides the same strong execution guarantees, performs better, and
costs significantly less than existing standalone orchestrators. Compared with
a state-of-the-art standalone orchestrator, AWS Step Functions, \name{} grants
applications greater flexibility to employ application-specific patterns and
optimizations, runs as much as 2x faster and up to 9x cheaper for a
representative set of applications.}