\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix2019_v3}
\include{common}

\newcommand{\name}{\texttt{unum}}

%-------------------------------------------------------------------
\begin{document}
%-------------------------------------------------------------------

\date{}

% make title bold and 14 pt font (Latex default is non-bold, 16 pt)
% \title{\Large \bf \name{}: A Library Workflow System for Serverless Applications}
% \title{\Large \bf \name{}: Purely Serverless Workflows with Distributed
% Orchestration}
\title{Doing More with Less \\ Orchestrating Serverless Applications
without an Orchestrator}

%for single author (just remove % characters)
\author{
  Paper \#224
% {\rm David H. Liu}\\
% Princeton University
% \and
% {\rm Amit Levy}\\
% Princeton University
% \and
% {\rm Shadi Noghabi}\\
% Microsoft Research
% \and
% {\rm Sebastian Burckhardt}\\
% Microsoft Research
% % copy the following lines to add more authors
% % \and
% % {\rm Name}\\
% %Name Institution
} % end author

\maketitle
\thispagestyle{empty}

\abstract
% Workflow systems have emerged to meet the demands of large serverless
% applications by providing high-level interfaces for workflow definition, rich
% semantics for function composition and strong guarantees for workflow
% execution.

Workflow systems have emerged to meet the demands of large-scale serverless
applications with higher-level programming interfaces, rich composition
semantics and strong execution guarantees. However, current designs rely on
supplemental hosted orchestrators to execute workflows and thus break out of
the serverless abstraction, compromising key serverless benefits such as fast
autoscaling, fine-grained billing and efficient resource multiplexing. \shadi{what is a "workflow" system? I am not sure it is a universaly well-know term...}

This paper presents \name{}, a new design for serverless workflow systems. 
In \name{}, a two-stage compiler distributes high-level orchestration
logic to constituent functions at compile-time. 
This removes the need for orchestrators at runtime, which allows workflows to 
execute purely as serverless functions. 
By checkpointing progress to a strongly consistent data store 
of choice, \name{} provides strong execution guarantees 
and minimizes re-executions under faults. \seb{re-execution is never completely avoidable.}

We present and evaluate an implementation of \name{} that can compile Step
Functions state machines and execute them purely as Lambda functions with
either S3 or DynamoDB as the data store. Our comparison with Step Functions
shows that applications scale up to 4.58x faster and costs more than one
order-of-magnitude less when running with \name{}. \shadi{you can shrink this last para to one sentence for the abstract.}


%-------------------------------------------------------------------
\input{introduction}
\input{background}
\input{design}
\input{impl}
\input{eval}
% \input{discussion}
\input{related}

%-------------------------------------------------------------------
\bibliographystyle{plain}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%  LocalWords:  endnotes includegraphics fread ptr nobj noindent
%%  LocalWords:  pdflatex acks
